---
title: "Visualizing COVID19 Mobility trends using google data"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
mob_trend <- readr::read_csv("global_mobility_trends.csv")
```


## Background

In this lab, we're going to use data visualizations to analyze and understand the influence of COVID19 on mobility trends using google data.

The goal of the lab is not to teach how to do the visualization. Don't worry, you will get there eventually, but it will take some more workshop meetings. 

The goal is to expose you to coding, let you play a little with code, and also show you some nice things that can be done.

In this lab you will have a chance to see:

   1. How a data structure looks like.
   2. How two different structures can have the same information, but serve different purposes.
   3. How data can be visualized.
   4. You will interpret and understand the visualization, using Israel's trends.
   5. You will change the code to make the visualization show a different country and try to interpret it as well.

## The data

Our data comes from google mobility trends, here: (https://www.google.com/covid19/mobility/)[https://www.google.com/covid19/mobility/].

They provide a baseline of "mobility" for each country, before the pandemic, and for every day during the pandemic. The data covers different areas such as residential, workplace, and more.

Run the following code to generate a visualization for the data:

```{r visualize-mobility-trends, exercise=TRUE}

mob_trend %>% 
  filter(country_region == "Israel") %>% 
  pivot_longer(cols = retail_and_recreation:residential, names_to = "area", values_to = "trend") %>% 
  ggplot(aes(x = date, y = trend, color = area)) + 
  geom_point(size = 0.5, alpha = 0.5) + 
  geom_smooth() +
  facet_wrap(~area)

```

Discuss the plot for a few minutes, try to interpret the different trends you see (for example in light of the first and second lockdowns we experienced). 

Answer these questions:

## Under the hood

## Critical thinking

## Summary and conclusions



